
---
title: "NSCLC2"
author: "Francesco Massaini"
date: "2024-03-28"
output: html_document
editor_options: 
  chunk_output_type: inline
---

# Imports
```{r Imports}
suppressMessages(library(dplyr)) 
suppressMessages(library(tibble))
suppressMessages(library(tidyr))
suppressMessages(library(org.Hs.eg.db)) # Homo sapiens db
suppressMessages(library(stringr))
suppressMessages(library(RColorBrewer)) # for plot
suppressMessages(library(pheatmap)) # for heatmaps
suppressMessages(library(ADImpute)) # This package provides functions to compute TPM
suppressMessages(library(ggplot2))
suppressMessages(library(FactoMineR))
suppressMessages(library(factoextra))
suppressMessages(library(survival))
suppressMessages(library(ranger))
suppressMessages(library(ggfortify))
suppressMessages(library(rlang))
suppressMessages(library(purrr))
suppressMessages(library(rio))
suppressMessages(library(uwot))
suppressMessages(library(DESeq2))
```


# Functions
```{r Functions}
source("../Functions.R")
```

# Data loading and processiong
## Load NSCLC2 data
```{r}
# Expression data
raw_counts = read.delim("../../IMMUCAN_data/NSCLC2/01_Raw_Counts/NSCLC2_all_patients_counts_forDE.txt", sep = "\t", check.names = TRUE) # "-" it's automatically converted into "." with checknames = TRUE (default). With FALSE, the script crashes. Let's convert . into original - manually
colnames(raw_counts) = str_replace_all(colnames(raw_counts), pattern = "\\.", replacement = "-")
 
  

# Clinical data
clinical = read.csv("../../IMMUCAN_data/NSCLC2/01_Clinical_Data/Daniel/NSCLC2_Daniel_relevant_columns.csv", row.names = 1) # Load
```

## making sure the row names in clinical matches to columns names in raw_counts and with the same order (DEseq requirements)
```{r}
colnames(raw_counts) = str_split_i(colnames(raw_counts), pattern = "-FIXT-.*", i = 1) # no issues with multiple samples for each patients because there is only one sample for each patients

clinical = clinical %>%
  filter(rownames(.) %in% colnames(raw_counts)) # select only common patients 

identical(rownames(clinical), colnames(raw_counts)) # identical checks also for the order

```

# pre-filtering: removing rows with low gene counts
```{r}
compute_distribution(data = raw_counts, plot_title = "Distribution of counts", xlab = "log2(Counts+1)", use_log = T) # log2(x+1)
filtered_raw_counts = filter_low_expr_genes(raw_counts, avg_expr_threshold = 12, zero_count_percent_threshold = 0.8)
compute_distribution(data = filtered_raw_counts, plot_title = "Distribution of counts", xlab = "log2(Counts+1)", use_log = T) # log2(x+1)
cat("Number of final genes:", nrow(filtered_raw_counts), "\nInitial number of genes:",  nrow(raw_counts), "\n Difference:", nrow(raw_counts)-nrow(filtered_raw_counts))
```



## Construct a DESeqDataSet object
```{r}
dds = DESeqDataSetFromMatrix(countData = filtered_raw_counts,
                       colData = clinical,
                       design = ~ death) # design specify the column of clinical data to use
dds
```
# set the factor level
```{r}
dds$death <- relevel(dds$death, ref = "Yes")
```

# Run DESeq on the previously created object
```{r}
dds <- DESeq(dds) # run DESeq and saving back results on the same object
res <- results(dds) # saving DESeq results in res

res
# results columns
# baseMean : average of normalized counts
# log2FoldChange : FC of genes compared between conditions     
# lfcSE : standard estimate of logFC 
# stat : wall test values for each genes
# pvalue : pvalue
# padj : padj for multiple testing


resLFC <- lfcShrink(dds, coef="death_No_vs_Yes", type="ashr")
```
# Explore Results 
```{r}
summary(res)
summary(resLFC)
cat("Number of adjusted p-values were less than 0.1: ", sum(res$padj < 0.1, na.rm=TRUE), "\n")
res0.01 <- results(dds, alpha = 0.01)
summary(res0.01)

```
# MA plot
```{r}
DESeq2::plotMA(res, ylim=c(-2,2))
DESeq2::plotMA(resLFC, ylim=c(-2,2))
# Blu genes are significant differential expressed genes
## NOTE: none genes in the top right corners (highly expressed - high mean normalized count - and high fold change)
```
## Over/Under expressed genes 
Let's select significant and highly expressed (high normalized mean count) genes and sort by LFC
```{r}
FC_threshold = 1
p_adj_threshold = 0.05
# Geni sovraespressi
overexpressed_genes <- resLFC[resLFC$log2FoldChange > FC_threshold & resLFC$padj < p_adj_threshold, ]
length(overexpressed_genes@rownames)
# Geni sottoespressi
underexpressed_genes <- resLFC[resLFC$log2FoldChange < -FC_threshold & resLFC$padj < p_adj_threshold, ]
length(underexpressed_genes@rownames)

```

```{r}
df_overexpressed = data.frame("genes" = overexpressed_genes@rownames, 
                              "base_mean" = overexpressed_genes@listData[["baseMean"]], 
                              "logFC" = overexpressed_genes@listData[["log2FoldChange"]], 
                              "p_adj" = overexpressed_genes@listData[["padj"]]) 
logFC_order = order(df_overexpressed$logFC, decreasing = TRUE)

df_overexpressed = df_overexpressed[logFC_order,]

df_downexpressed = data.frame("genes" = underexpressed_genes@rownames, 
                              "base_mean" = underexpressed_genes@listData[["baseMean"]], 
                              "logFC" = underexpressed_genes@listData[["log2FoldChange"]], 
                              "p_adj" = underexpressed_genes@listData[["padj"]]) 
logFC_order = order(df_downexpressed$logFC, decreasing = TRUE)

df_downexpressed = df_downexpressed[logFC_order,]

```


```{r}
ov_counts_df = filtered_raw_counts[df_overexpressed$genes,]
ov_counts_df = log2(ov_counts_df+1)

un_counts_df = filtered_raw_counts[df_downexpressed$genes,]
un_counts_df = log2(un_counts_df+1)

DE_counts_df = rbind(ov_counts_df, un_counts_df)
```


```{r}
Compute_Heatmaps(expr_df = DE_counts_df, sample_metadata_df = clinical)
```

## PCA biplot
```{r}
# PCA biplot on TPM
pca_result <- FactoMineR::PCA(t(DE_counts_df), scale.unit = TRUE, graph = FALSE)  
fviz_eig(pca_result) # Scree plot: explained variance by PC

# pdf(paste0(output_folder_TPM_outlier, "TPM_samples_PCA.pdf"))
for (feat in colnames(clinical)){
  p = factoextra::fviz_pca_biplot(pca_result, repel = T, select.var = list(contrib = 5), label = "var",
                              habillage = as.factor(clinical[,feat]),
                              addEllipses = T, ellipse.level=0.95, palette = "Dark2", 
                              legend.title = feat,
                              ggtheme = theme_minimal())
  print(p)
}
# repel = T do not overlap the text, select.var = list(contrib = 5) shows only the most 5 variable feature (variables) that contributes to separation. label = var , put label ONLY on the 5 variables
# dev.off()
```
# Same on multiple factors
```{r}
clinical[,"death"] <- as.factor(clinical[,"death"])
clinical[,"adjuvant_treatment"] <- as.factor(clinical[,"adjuvant_treatment"])

ddsMF = DESeqDataSetFromMatrix(countData = filtered_raw_counts,
                       colData = clinical,
                       design = ~ death + adjuvant_treatment) # design specify the column of clinical data to use
 


ddsMF <- DESeq(ddsMF)

resMFType <- results(ddsMF,
                     contrast=c("adjuvant_treatment", "Yes", "No"))
```

Let's select significant and highly expressed (high normalized mean count) genes and sort by LFC
```{r}
FC_threshold = 1
p_adj_threshold = 0.05
# Geni sovraespressi
overexpressed_genes <- resMFType[resMFType$log2FoldChange > FC_threshold & resMFType$padj < p_adj_threshold, ]
length(overexpressed_genes@rownames)
# Geni sottoespressi
underexpressed_genes <- resMFType[resMFType$log2FoldChange < -FC_threshold & resMFType$padj < p_adj_threshold, ]
length(underexpressed_genes@rownames)

```

```{r}
df_overexpressed = data.frame("genes" = overexpressed_genes@rownames, 
                              "base_mean" = overexpressed_genes@listData[["baseMean"]], 
                              "logFC" = overexpressed_genes@listData[["log2FoldChange"]], 
                              "p_adj" = overexpressed_genes@listData[["padj"]]) 
logFC_order = order(df_overexpressed$logFC, decreasing = TRUE)

df_overexpressed = df_overexpressed[logFC_order,]

df_downexpressed = data.frame("genes" = underexpressed_genes@rownames, 
                              "base_mean" = underexpressed_genes@listData[["baseMean"]], 
                              "logFC" = underexpressed_genes@listData[["log2FoldChange"]], 
                              "p_adj" = underexpressed_genes@listData[["padj"]]) 
logFC_order = order(df_downexpressed$logFC, decreasing = TRUE)

df_downexpressed = df_downexpressed[logFC_order,]

```


```{r}
ov_counts_df = filtered_raw_counts[df_overexpressed$genes,]
ov_counts_df = log2(ov_counts_df+1)

un_counts_df = filtered_raw_counts[df_downexpressed$genes,]
un_counts_df = log2(un_counts_df+1)

DE_counts_df = rbind(ov_counts_df, un_counts_df)
```


```{r}
Compute_Heatmaps(expr_df = DE_counts_df, sample_metadata_df = clinical)
```
## PCA biplot
```{r}
# PCA biplot on TPM
pca_result <- FactoMineR::PCA(t(DE_counts_df), scale.unit = TRUE, graph = FALSE)  
fviz_eig(pca_result) # Scree plot: explained variance by PC

# pdf(paste0(output_folder_TPM_outlier, "TPM_samples_PCA.pdf"))
for (feat in colnames(clinical)){
  p = factoextra::fviz_pca_biplot(pca_result, repel = T, select.var = list(contrib = 5), label = "var",
                              habillage = as.factor(clinical[,feat]),
                              addEllipses = T, ellipse.level=0.95, palette = "Dark2", 
                              legend.title = feat,
                              ggtheme = theme_minimal())
  print(p)
}
# repel = T do not overlap the text, select.var = list(contrib = 5) shows only the most 5 variable feature (variables) that contributes to separation. label = var , put label ONLY on the 5 variables
# dev.off()
```




